---
title: "When and Why Advice-Taking Strategies Differ:"
subtitle: "A Dual-Hurdle-Based Meta-Analysis"
format: 
  revealjs:
    css: styles.css
    show-slide-number: all
width: 1200
author: "Jessica Helmer | M^3^ | 2026-06-23"
html-table-processing: none
self-contained: true
---

## Advice-Taking

```{r}
#| include: false
library(tidyverse)
library(targets)
library(gt)
library(patchwork)

tar_load(c(study_counts, dat, 
           ppc_plt, ppc_plt_g_subsample, 
           itemlvl_plt, personlvl_plt, studylvl_plt), 
         store = here::here("_targets"))

patched_pred_plt_files <- tar_read(patched_pred_plt_files, store = here::here("_targets")) |>
  unlist() |>
  normalizePath(winslash = "/") |> 
  here::here()

# print(patched_pred_plt_files)
# dat <- readRDS(here::here("..", "Data Clean", "trihurdle-metaanalysis-more-predictors_dat-small_2026-03.15.rds"))
# study_names <- readRDS(here::here("..", "Data Clean", "studynames.rds"))
# condition_names <- readRDS(here::here("..", "Data Clean", "condition-names_more-predictors.rds"))
# data_study_counts <- readRDS(here::here("..", "Data Clean", "data_study_counts.rds"))

main_color <- "steelblue"
golden_ratio <- (1 + sqrt(5)) / 2

set_theme(theme_classic(base_size = 24))
theme_update(aspect.ratio = golden_ratio ^ -1,
             axis.line.y = element_blank(),
             axis.ticks.y = element_blank(),
             axis.text.x = element_text(size = 30))

# study.level_predictors <- str_split(condition_names[1], regex("\\.[lmh]{1}_")) |> 
#   map(~ str_split_i(.x, "\\.", 1)) |>
#   pluck(1) %>%
#   tail(5)
```

Studies on advice taking often use a Judge Advisor System setup

<div class="subpoint">○ Judge makes initial judgement</div>

<div class="subpoint">○ Judge receives advice</div>

<div class="subpoint">○ Judge makes final judgement</div>

Outcome metric commonly "Weight of Advice" (WOA)

<!-- Outcome metric of interest is how much someone changed their prediction relative to the advice -->

<br>

:::: {.columns}
::: {.column width="50%"}
$$ WOA = \frac{\text{initial} - \text{final}}{\text{initial} - \text{advice}} $$
:::

::: {.column width="50%"}
```{r}
#| layout-align: center
#| layout-valign: center
tibble(woa = c("$WOA = 0$", "$WOA = 1$", "$0 < WOA < 1$"),
        res = c("Decline", "Adopt", "Compromise")) |>
  gt() |>
  fmt_markdown() |>
  tab_options(column_labels.hidden = T) |>
  opt_table_font(size = 30) |>
  opt_horizontal_padding(3)
```
:::
::::


## Weight of Advice

WOA continuous metric, but often empirically see concentrations of values at 0 and 1

```{r}
#| fig-align: center
#| fig-height: 4
dat |>
  filter(study %in% c(1, 24, 28)) |>
  mutate(study = as.character(study)) |>
  # left_join(study_names, by = "study") |>
  mutate(studyname = str_replace_all(studyname, "_", " ")) |>
  ggplot(aes(x = woa_winsor, y = after_stat(density))) +
  geom_histogram(color = NA, fill = main_color, binwidth = 0.08) +
  scale_x_continuous(breaks = c(0, 0.5, 1), labels = c("0", ".5", "1")) +
  scale_y_continuous(limits = c(0, 6.7), breaks = c(0, 6.7), labels = NULL) +
  guides(x = guide_axis(cap = "both")) +
  labs(y = NULL, x = "WOA") +
  facet_wrap(~ studyname, scales = "free_y") +
  theme(panel.spacing = unit(40, "pt"),
        axis.text.x = element_text(size = 20),
        plot.margin = unit(c(-10, 0, -10, 0), "pt"))
```

Typically are modeled with standard linear regression procedures

<div class="subpoint">○ Assume normally distributed residuals!</div>

::: footer
(Himmelstein, 2022)
:::


## Dual-Hurdle Model

Separates outcome into two phases:

:::: {.columns}
::: {.column width="50%"}
**Discrete**

::: {style="font-size:90%"}
Decline, adopt, or compromise?

Categorical logit: estimates probability of each choice
:::

```{r}
tibble(choice = factor(c("Decline", "Compromise", "Adopt"), 
                       levels = c("Decline", "Compromise", "Adopt"), ordered = T),
       p = c(0.45, 0.25, 0.30)) |>
  ggplot(aes(x = choice, y = p)) +
  geom_col(fill = main_color) +
  scale_y_continuous(NULL, limits = c(0, .46), breaks = NULL) +
  guides(x = guide_axis(NULL, cap = "both"))
```
:::

::: {.column width="50%"}
**Continuous**

::: {style="font-size:90%"}
If compromise, how much?

Beta distribution: estimates continuous WOA between 0 and 1
:::

```{r}
tibble(woa = rbeta(10000, 1.5, 3)) |>
  ggplot(aes(x = woa)) +
  geom_density(fill = main_color, color = NA, adjust = 3) +
  scale_y_continuous(NULL, limits = c(0, 1.8), breaks = NULL) +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1), labels = c("0", ".5", "1")) +
  guides(x = guide_axis(NULL, cap = "both")) 
```
:::
::::

::: footer
(Himmelstein, 2022)
:::


## Dual-Hurdle Model

:::: {.columns}
::: {.column width="55%"}
Makes WOA distributions like...

```{r}
#| fig-height: 8
d <- rbind(tibble(pc1 = .25,
        pc2 = .45,
        pc3 = .30,
        a = 1.5,
        b = 3) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 1),
   tibble(pc1 = .4,
        pc2 = .5,
        pc3 = .1,
        a = 2,
        b = 4) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 2),
   tibble(pc1 = .7,
        pc2 = .15,
        pc3 = .15,
        a = 3,
        b = 4) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 3),
   tibble(pc1 = .2,
        pc2 = .2,
        pc3 = .6,
        a = 5,
        b = 6) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 4))

d |>
  ggplot(aes(x = woa)) +
  geom_density(fill = main_color, color = NA, adjust = 1.1) +
  scale_y_continuous(NULL, breaks = NULL) +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1),
                     labels = c("0", ".5", "1")) +
  facet_wrap(~ version, ncol = 2, scales = "free_y", axes = "all_x") + 
  guides(x = guide_axis("WOA", cap = "both")) +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 25),
        plot.margin = unit(rep(-1, 4), "pt"),
        strip.background = element_blank(),
        strip.text.x = element_blank())
```
:::

::: {.column width="45%"}
Can examine effects of predictors for both the categorical and continuous pieces

Captures variation at the item- and person-level

Present study: extend dual-hurdle model to meta-analytic context

<div class="subpoint">○ Account for study-level variation</div>
:::
::::

::: footer
(Himmelstein, 2022)
:::


## Present Study

Used data collected for an advice-taking metaanalysis on gender effects (Kausel et al., 2024)

<div class="subpoint">○ Full dataset included `r study_counts |> filter(version == "full") |> pull(studies)` studies and `r study_counts |> filter(version == "full") |> pull(observations) |> format(big.mark = ",")` observations.</div>

Removed cases that had missing data for WOA or any of the predictors

<div class="subpoint">○ Remaining data had `r study_counts |> filter(version == "no missing") |> pull(studies)` studies: `r study_counts |> filter(version == "no missing") |> pull(observations) |> format(big.mark = ",")`.</div>

```{r}
tibble("Trial-Level" = list(list("advice distance")),
       "Person-Level" = list(list("age", "gender")),
       "Study-Level" = list(list("student population", "almanac item", "future item", "expert advisor", "incentive"))) |>
  mutate(across(everything(), ~ map_chr(.x, ~ paste(.x, collapse = "<br>○ ")) |>
                  paste0("○ ", ... = _) |> str_to_title())) |>
  gt() |>
  fmt_markdown() |>
  tab_style(cell_text(weight = "bold", align = "center"),
             cells_column_labels()) |>
  tab_style(cell_text(v_align = "top"),
            cells_body()) |>
  tab_header("Predictors") |>
  opt_table_font(size = 22) |>
  opt_horizontal_padding(1.5)
```

::: footer
(Kausel et al., 2024)
:::


## Dual-Hurdle Model

Distribution of WOA values as per the **data** and the [model predictions]{style="color:`r main_color`"}

```{r}
#| fig-align: center
#| message: false
p <- readRDS(here::here("figures", "metaanalysis", "ppc_plt.rds")) +
  theme_classic(base_size = 24) +
  scale_color_manual(values = c("yrep" = main_color, "y" = "black")) +
  scale_y_continuous(NULL, breaks = NULL)  +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1),
                     labels = c("0", ".5", "1")) + 
  guides(x = guide_axis("WOA", cap = "both")) +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 25),
        plot.margin = unit(rep(-1, 4), "pt"),
        legend.position = "none",
        aspect.ratio = golden_ratio ^ -1)

p@layers$stat_density$aes_params$linewidth <- p@layers$stat_density$aes_params$linewidth * 2.5
p@layers$stat_density...2$aes_params$linewidth <- p@layers$stat_density...2$aes_params$linewidth * 2

p
```


## Dual-Hurdle Model

Distribution of WOA values as per the **data** and the [model predictions]{style="color:`r main_color`"}

```{r}
#| fig-align: center
#| message: false
p +
  annotate("text", label = "Surplus of WOA values at exactly WOA = .5!" |> str_wrap(20),
           x = 0.625, y = 2.6, hjust = 0, size = 7) +
  annotate(geom = "curve", x = 0.61, y = 2.6, xend = 0.5, yend = 1.3, 
    curvature = .4, arrow = arrow(length = unit(2, "mm")))
```



<!-- ## Tri-Hurdle Model -->

<!-- Added a categorical choice: Decline, Adopt, **Midpoint**, Compromise -->

<!-- ```{r} -->
<!-- #| fig-align: center -->
<!-- #| message: false -->
<!-- # p <- readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis", "ppc_plt.rds")) + -->
<!-- #   theme_classic(base_size = 24) + -->
<!-- #   scale_color_manual(values = c("yrep" = main_color, "y" = "black")) + -->
<!-- #   scale_y_continuous(NULL, breaks = NULL)  + -->
<!-- #   scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1), -->
<!-- #                      labels = c("0", ".5", "1")) +  -->
<!-- #   guides(x = guide_axis("WOA", cap = "both")) + -->
<!-- #   theme(axis.line.y = element_blank(), -->
<!-- #         axis.ticks.y = element_blank(), -->
<!-- #         axis.text.x = element_text(size = 25), -->
<!-- #         plot.margin = unit(rep(-1, 4), "pt"), -->
<!-- #         legend.position = "none", -->
<!-- #         aspect.ratio = golden_ratio ^ -1) -->
<!-- #  -->
<!-- # p@layers$stat_density$aes_params$linewidth <- p@layers$stat_density$aes_params$linewidth * 2.5 -->
<!-- # p@layers$stat_density...2$aes_params$linewidth <- p@layers$stat_density...2$aes_params$linewidth * 2 -->
<!-- #  -->
<!-- # p -->

<!-- p <- ppc_plt + -->
<!--   theme_classic(base_size = 24) + -->
<!--   scale_color_manual(values = c("yrep" = main_color, "y" = "black")) + -->
<!--   scale_y_continuous(NULL, breaks = NULL)  + -->
<!--   scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1), -->
<!--                      labels = c("0", ".5", "1")) +  -->
<!--   guides(x = guide_axis("WOA", cap = "both")) + -->
<!--   theme(axis.line.y = element_blank(), -->
<!--         axis.ticks.y = element_blank(), -->
<!--         axis.text.x = element_text(size = 25), -->
<!--         plot.margin = unit(rep(-1, 4), "pt"), -->
<!--         legend.position = "none", -->
<!--         aspect.ratio = golden_ratio ^ -1) -->

<!-- p@layers$stat_density$aes_params$linewidth <- p@layers$stat_density$aes_params$linewidth * 2.5 -->
<!-- p@layers$stat_density...2$aes_params$linewidth <- p@layers$stat_density...2$aes_params$linewidth * 2 -->

<!-- p -->
<!-- ``` -->


<!-- ## Tri-Hurdle Model -->

<!-- ```{r} -->
<!-- #| fig-align: center -->
<!-- #| message: false -->
<!-- # p <- readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis", "ppc_plt_g.rds")) + -->
<!-- #  theme_classic() + -->
<!-- #  scale_color_manual(values = c("yrep" = main_color, "y" = "black")) + -->
<!-- #  scale_y_continuous(NULL, breaks = NULL) + -->
<!-- #  scale_x_continuous(NULL, limits = c(0, 1), breaks = c(0, .5, 1), -->
<!-- #                     labels = NULL) + -->
<!-- #  guides(x = guide_axis(cap = "both")) + -->
<!-- #  theme(axis.line.y = element_line(color = NA), -->
<!-- #        axis.ticks.y = element_blank(), -->
<!-- #        #axis.text.x = element_text(size = 4), -->
<!-- #        # strip.text = element_text(size = 6), -->
<!-- #        # strip.background = element_rect(color = NA), -->
<!-- #        strip.text = element_blank(), -->
<!-- #        strip.background = element_blank(), -->
<!-- #        #aspect.ratio = golden_ratio ^ -1, -->
<!-- #        legend.position = "none") -->
<!-- #  -->
<!-- # p@layers$stat_density$aes_params$linewidth <- p@layers$stat_density$aes_params$linewidth * 2.5 -->
<!-- # p@layers$stat_density...2$aes_params$linewidth <- p@layers$stat_density...2$aes_params$linewidth * 0.7 -->
<!-- #  -->
<!-- # p -->

<!-- p <- ppc_plt_g_subsample + -->
<!--  scale_color_manual(values = c("yrep" = main_color, "y" = "black")) + -->
<!--  scale_y_continuous(NULL, breaks = NULL) + -->
<!--  scale_x_continuous(NULL, limits = c(0, 1), breaks = c(0, .5, 1), -->
<!--                     labels = NULL) + -->
<!--  guides(x = guide_axis(cap = "both")) + -->
<!--  theme(axis.line.y = element_line(color = NA), -->
<!--        axis.ticks.y = element_blank(), -->
<!--        #axis.text.x = element_text(size = 4), -->
<!--        # strip.text = element_text(size = 6), -->
<!--        # strip.background = element_rect(color = NA), -->
<!--        strip.text = element_blank(), -->
<!--        strip.background = element_blank(), -->
<!--        #aspect.ratio = golden_ratio ^ -1, -->
<!--        legend.position = "none") -->

<!-- p@layers$stat_density$aes_params$linewidth <- p@layers$stat_density$aes_params$linewidth * 2.5 -->
<!-- p@layers$stat_density...2$aes_params$linewidth <- p@layers$stat_density...2$aes_params$linewidth * 0.7 -->

<!-- p -->
<!-- ``` -->


## Decline, Adopt, Midpoint, Compromise {data-center="true"}

```{r}
#| fig-width: 4
#| fig-height: 2.5
#| fig-align: center
#| layout-ncol: 3

personlvl_plt
itemlvl_plt
studylvl_plt
```

## Distance

```{r}
#| fig-width: 6.5
#| fig-align: center
patched_pred_plt_files[[1]] |> knitr::include_graphics(error = F)
```


## Age and Gender

```{r}
#| fig-width: 6.5
#| fig-align: center
#| layout-ncol: 2
patched_pred_plt_files[[2]] |> knitr::include_graphics(error = F)
patched_pred_plt_files[[3]] |> knitr::include_graphics(error = F)
```


## Study-Level Predictors

```{r}
#| fig-width: 6.5
#| fig-align: center
#| layout-ncol: 2
patched_pred_plt_files[[4]] |> knitr::include_graphics(error = F)
patched_pred_plt_files[[5]] |> knitr::include_graphics(error = F)
```

## Study-Level Predictors

```{r}
#| fig-width: 6.5
#| fig-align: center
#| layout-ncol: 2
patched_pred_plt_files[[6]] |> knitr::include_graphics(error = F)
patched_pred_plt_files[[7]] |> knitr::include_graphics(error = F)
```

## Study-Level Predictors

```{r}
#| fig-width: 6.5
#| fig-align: center
patched_pred_plt_files[[8]] |> knitr::include_graphics(error = F)
```


## Discussion

Midpoint averaging common, could be relevant to include in model

<div class="subpoint">○ Varies by study</div>

Differences between student and non-student populations

<div class="subpoint">○ Declining a lot, but updating more when compromise</div>

Differences between incentivised and non-incentivised studies


## References

::: {style="font-size:70%"}
Kausel, E.E., Ventura, S., Reyes, V. (November 2024). [Conference presentation]. Gender Differences in Advice Taking: A Meta-Analytic Path Model. Society of Judgement and Decision Making, New York, NY.

Himmelstein, M. (2022). Decline, adopt or compromise? A dual hurdle model for advice utilization. Journal of Mathematical Psychology, 110, 102695. https://doi.org/10.1016/j.jmp.2022.102695

<br>

This research was supported in part through research cyberinfrastructure resources and services provided by the [**Partnership for an Advanced Computing Environment (PACE)**]{style="color:#857437"} at the [**Georgia Institute of Technology**]{style="color:#003057"}, Atlanta, Georgia, USA. RRID:SCR_027619
:::

# Thank You!

Contact: jhelmer3@gatech.edu

```{r}
plot(qrcode::qr_code("https://jhelmer3.github.io/advice-taking-metaanalysis/presentations/M3_2026.html#/title-slide"))
```



```{r}
#| fig-align: center
#| fig-height: 7
#| fig-width: 13
#| eval: false

top_plts <- study.level_predictors |>
  head(3) |>
  map(~ (readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis More Predictors",
                           paste0(.x, "_plt_small.rds"))) & 
            theme(geom = element_geom(ink = "#597993", paper = "#a2b9cd"))))

bottom_plts <- study.level_predictors |>
  tail(2) |>
  map(~ (readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis More Predictors",
                           paste0(.x, "_plt_small.rds"))) & 
            theme(geom = element_geom(ink = "#597993", paper = "#a2b9cd"))))

wrap_plots(top_plts, ncol = 3) / 
  wrap_plots(plot_spacer(), wrap_plots(bottom_plts, ncol = 2), plot_spacer(),
             widths = c(0.25, 1, 0.25))
```




