---
title: "When and Why Advice-Taking Strategies Differ:"
subtitle: "A Dual-Hurdle-Based Meta-Analysis"
format: 
  revealjs:
    css: styles.css
    show-slide-number: all
width: 1200
author: "Jessica Helmer | TeaP | 2026-03-16"
html-table-processing: none
self-contained: true
---

## Advice-Taking

```{r}
#| include: false
library(tidyverse)
library(gt)
library(patchwork)

dat <- readRDS(here::here("..", "Data Clean", "trihurdle-metaanalysis-more-predictors_dat-small_2026-03.15.rds"))
study_names <- readRDS(here::here("..", "Data Clean", "studynames.rds"))
condition_names <- readRDS(here::here("..", "Data Clean", "condition-names_more-predictors.rds"))
data_study_counts <- readRDS(here::here("..", "Data Clean", "data_study_counts.rds"))

main_color <- "steelblue"
golden_ratio <- (1 + sqrt(5)) / 2

set_theme(theme_classic(base_size = 24))
theme_update(aspect.ratio = golden_ratio ^ -1,
             axis.line.y = element_blank(),
             axis.ticks.y = element_blank(),
             axis.text.x = element_text(size = 30))

study.level_predictors <- str_split(condition_names[1], regex("\\.[lmh]{1}_")) |> 
  map(~ str_split_i(.x, "\\.", 1)) |>
  pluck(1) %>%
  tail(5)
```

Studies on advice taking often use a Judge Advisor System setup

<div class="subpoint">○ Judge makes initial judgement</div>

<div class="subpoint">○ Judge receives advice</div>

<div class="subpoint">○ Judge makes final judgement</div>

Outcome metric commonly "Weight of Advice" (WOA)

<!-- Outcome metric of interest is how much someone changed their prediction relative to the advice -->

<br>

:::: {.columns}
::: {.column width="50%"}
$$ WOA = \frac{\text{initial} - \text{final}}{\text{initial} - \text{advice}} $$
:::

::: {.column width="50%"}
```{r}
#| layout-align: center
#| layout-valign: center
tibble(woa = c("$WOA = 0$", "$WOA = 1$", "$0 < WOA < 1$"),
        res = c("Decline", "Adopt", "Compromise")) |>
  gt() |>
  fmt_markdown() |>
  tab_options(column_labels.hidden = T) |>
  opt_table_font(size = 30) |>
  opt_horizontal_padding(3)
```
:::
::::


## Weight of Advice

WOA continuous metric, but often empirically see concentrations of values at 0 and 1

```{r}
#| fig-align: center
#| fig-height: 4
dat |>
  filter(study %in% c(1, 4, 11)) |>
  mutate(study = as.character(study)) |>
  # left_join(study_names, by = "study") |>
  mutate(studyname = str_replace_all(studyname, "_", " ")) |>
  ggplot(aes(x = woa_winsor, y = after_stat(density))) +
  geom_histogram(color = NA, fill = main_color, binwidth = 0.08) +
  scale_x_continuous(breaks = c(0, 0.5, 1), labels = c("0", ".5", "1")) +
  scale_y_continuous(limits = c(0, 6.7), breaks = c(0, 6.7), labels = NULL) +
  guides(x = guide_axis(cap = "both")) +
  labs(y = NULL, x = "WOA") +
  facet_wrap(~ studyname, scales = "free_y") +
  theme(panel.spacing = unit(40, "pt"),
        axis.text.x = element_text(size = 20),
        plot.margin = unit(c(-10, 0, -10, 0), "pt"))
```

Typically are modeled with standard linear regression procedures

<div class="subpoint">○ Assume normally distributed residuals!</div>

::: footer
(Himmelstein, 2022)
:::


## Dual-Hurdle Model

Separates outcome into two phases:

:::: {.columns}
::: {.column width="50%"}
**Discrete**

::: {style="font-size:90%"}
Decline, adopt, or compromise?

Categorical logit: estimates probability of each choice
:::

```{r}
tibble(choice = factor(c("Decline", "Compromise", "Adopt"), 
                       levels = c("Decline", "Compromise", "Adopt"), ordered = T),
       p = c(0.45, 0.25, 0.30)) |>
  ggplot(aes(x = choice, y = p)) +
  geom_col(fill = main_color) +
  scale_y_continuous(NULL, limits = c(0, .46), breaks = NULL) +
  guides(x = guide_axis(NULL, cap = "both"))
```
:::

::: {.column width="50%"}
**Continuous**

::: {style="font-size:90%"}
If compromise, how much?

Beta distribution: estimates continuous WOA between 0 and 1
:::

```{r}
tibble(woa = rbeta(10000, 1.5, 3)) |>
  ggplot(aes(x = woa)) +
  geom_density(fill = main_color, color = NA, adjust = 3) +
  scale_y_continuous(NULL, limits = c(0, 1.8), breaks = NULL) +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1), labels = c("0", ".5", "1")) +
  guides(x = guide_axis(NULL, cap = "both")) 
```
:::
::::

::: footer
(Himmelstein, 2022)
:::


## Dual-Hurdle Model

:::: {.columns}
::: {.column width="55%"}
Makes WOA distributions like...

```{r}
#| fig-height: 8
d <- rbind(tibble(pc1 = .25,
        pc2 = .45,
        pc3 = .30,
        a = 1.5,
        b = 3) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 1),
   tibble(pc1 = .4,
        pc2 = .5,
        pc3 = .1,
        a = 2,
        b = 4) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 2),
   tibble(pc1 = .7,
        pc2 = .15,
        pc3 = .15,
        a = 3,
        b = 4) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 3),
   tibble(pc1 = .2,
        pc2 = .2,
        pc3 = .6,
        a = 5,
        b = 6) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 4))

d |>
  ggplot(aes(x = woa)) +
  geom_density(fill = main_color, color = NA, adjust = 1.1) +
  scale_y_continuous(NULL, breaks = NULL) +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1),
                     labels = c("0", ".5", "1")) +
  facet_wrap(~ version, ncol = 2, scales = "free_y", axes = "all_x") + 
  guides(x = guide_axis("WOA", cap = "both")) +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 25),
        plot.margin = unit(rep(-1, 4), "pt"),
        strip.background = element_blank(),
        strip.text.x = element_blank())
```
:::

::: {.column width="45%"}
Can examine effects of predictors for both the categorical and continuous pieces

Captures variation at the item- and person-level

Present study: extend dual-hurdle model to meta-analytic context

<div class="subpoint">○ Account for study-level variation</div>
:::
::::

::: footer
(Himmelstein, 2022)
:::


## Present Study

Used data collected for an advice-taking metaanalysis on gender effects (Kausel et al., 2024)

<div class="subpoint">○ Full dataset included `r data_study_counts |> filter(version == "full") |> pull(studies)` studies and `r data_study_counts |> filter(version == "full") |> pull(observations) |> format(big.mark = ",")` observations.</div>

Removed cases that had missing data for WOA or any of the predictors

<div class="subpoint">○ Remaining data had `r data_study_counts |> filter(version == "no missing") |> pull(studies)` studies: `r data_study_counts |> filter(version == "no missing") |> pull(observations) |> format(big.mark = ",")`.</div>

```{r}
tibble("Trial-Level" = list(list("advice distance", "confidence")),
       "Person-Level" = list(list("age", "gender")),
       "Study-Level" = list(list("student population", "almanac item", "future item", "expert advisor", "incentive"))) |>
  mutate(across(everything(), ~ map_chr(.x, ~ paste(.x, collapse = "<br>○ ")) |>
                  paste0("○ ", ... = _) |> str_to_title())) |>
  gt() |>
  fmt_markdown() |>
  tab_style(cell_text(weight = "bold", align = "center"),
             cells_column_labels()) |>
  tab_style(cell_text(v_align = "top"),
            cells_body()) |>
  tab_header("Predictors") |>
  opt_table_font(size = 22) |>
  opt_horizontal_padding(1.5)
```

::: footer
(Kausel et al., 2024)
:::


## Unconditional Model

Distribution of WOA values as per the **data** and the [model predictions]{style="color:`r main_color`"}

```{r}
#| fig-align: center
#| message: false
p <- readRDS(here::here("Figures", "Metaanalysis", "ppc_plt.rds")) +
  theme_classic(base_size = 24) +
  scale_color_manual(values = c("yrep" = main_color, "y" = "black")) +
  scale_y_continuous(NULL, breaks = NULL)  +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1),
                     labels = c("0", ".5", "1")) + 
  guides(x = guide_axis("WOA", cap = "both")) +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 25),
        plot.margin = unit(rep(-1, 4), "pt"),
        legend.position = "none",
        aspect.ratio = golden_ratio ^ -1)

p@layers$stat_density$aes_params$linewidth <- p@layers$stat_density$aes_params$linewidth * 2.5
p@layers$stat_density...2$aes_params$linewidth <- p@layers$stat_density...2$aes_params$linewidth * 2

p
```


## Dual-Hurdle Model

Distribution of WOA values as per the **data** and the [model predictions]{style="color:`r main_color`"}

```{r}
#| fig-align: center
#| message: false
p +
  annotate("text", label = "Surplus of WOA values at exactly WOA = .5!" |> str_wrap(20),
           x = 0.625, y = 2.6, hjust = 0, size = 7) +
  annotate(geom = "curve", x = 0.61, y = 2.6, xend = 0.5, yend = 1.3, 
    curvature = .4, arrow = arrow(length = unit(2, "mm")))
```



## Tri-Hurdle Model

Added a categorical choice: Decline, Adopt, **Midpoint**, Compromise

```{r}
#| fig-align: center
#| message: false
p <- readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis", "ppc_plt.rds")) +
  theme_classic(base_size = 24) +
  scale_color_manual(values = c("yrep" = main_color, "y" = "black")) +
  scale_y_continuous(NULL, breaks = NULL)  +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1),
                     labels = c("0", ".5", "1")) + 
  guides(x = guide_axis("WOA", cap = "both")) +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 25),
        plot.margin = unit(rep(-1, 4), "pt"),
        legend.position = "none",
        aspect.ratio = golden_ratio ^ -1)

p@layers$stat_density$aes_params$linewidth <- p@layers$stat_density$aes_params$linewidth * 2.5
p@layers$stat_density...2$aes_params$linewidth <- p@layers$stat_density...2$aes_params$linewidth * 2

p
```


## Tri-Hurdle Model

```{r}
#| fig-align: center
#| message: false
p <- readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis", "ppc_plt_g.rds")) +
  theme_classic() +
  scale_color_manual(values = c("yrep" = main_color, "y" = "black")) +
  scale_y_continuous(NULL, breaks = NULL) +
  scale_x_continuous(NULL, limits = c(0, 1), breaks = c(0, .5, 1),
                     labels = NULL) +
  guides(x = guide_axis(cap = "both")) +
  theme(axis.line.y = element_line(color = NA),
        axis.ticks.y = element_blank(),
        #axis.text.x = element_text(size = 4),
        # strip.text = element_text(size = 6),
        # strip.background = element_rect(color = NA),
        strip.text = element_blank(),
        strip.background = element_blank(),
        #aspect.ratio = golden_ratio ^ -1,
        legend.position = "none")

# p@layers$stat_density$aes_params$linewidth <- p@layers$stat_density$aes_params$linewidth * 2.5
p@layers$stat_density...2$aes_params$linewidth <- p@layers$stat_density...2$aes_params$linewidth * 0.7

p
```


## Decline, Adopt, Midpoint, Compromise

```{r}
#| fig-height: 3
#| fig-width: 4
#| fig-align: center

readRDS(here::here("..", "Data Clean", "trihurdle-metaanalysis_stdylvl_dat.rds")) |>
  ggplot(aes(x = Parameter, y = round(P, 2), fill = Parameter, color = Parameter)) + 
  geom_violin(color = NA, alpha = .6) + 
  geom_jitter(shape = 16, alpha = .8, width = .2) +
  theme_minimal() +
  xlab("Intercept") +
  ylab("Probability") +
  guides(fill = "none", color = "none") +
  scale_color_manual(NULL, values = c("#003f5c", "#7a5195", "#ef5675", "#ffa600")) +
  scale_fill_manual(NULL, values = c("#003f5c", "#7a5195", "#ef5675", "#ffa600")) +
  coord_cartesian(ylim = c(0, 1), expand = 0, clip = "off") +
  theme_minimal() +
  theme(panel.grid.minor = element_blank(),
        panel.grid.major.x = element_blank(),
        legend.position = "none",
        plot.margin = unit(c(30, 10, 30, 10), "pt"))
```


## Distance and Confidence 

```{r}
#| fig-width: 6.5
#| fig-align: center
readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis More Predictors", "distance-by-confidence_plt_small.rds"))
```


## Age and Gender

```{r}
#| fig-width: 6.5
#| fig-align: center
readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis More Predictors", "age-by-female_plt_small.rds"))
```


## Study-Level Predictors

```{r}
#| fig-align: center
#| fig-width: 13
study.level_predictors |>
  head(3) |>
  map(~ (readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis More Predictors",
                           paste0(.x, "_plt_small.rds"))) & 
            theme(geom = element_geom(ink = "#597993", paper = "#a2b9cd")) &)) |>
  wrap_plots(ncol = 3)
```


## Study-Level Predictors

```{r}
#| fig-align: center
#| fig-width: 13
#| out-width: 66%
plts <- study.level_predictors |>
  tail(2) |>
  map(~ (readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis More Predictors",
                           paste0(.x, "_plt_small.rds"))) & 
            theme(geom = element_geom(ink = "#597993", paper = "#a2b9cd")))) |>
  wrap_plots(bottom_plts, ncol = 2)

wrap_plots(plot_spacer(), plts, plot_spacer(),
             widths = c(0.25, 1, 0.25))
```


## Discussion

Midpoint averaging common, could be relevant to include in model

<div class="subpoint">○ Varies by study</div>

Differences between student and non-student populations


## References

Kausel, E.E., Ventura, S., Reyes, V. (November 2024). [Conference presentation]. Gender Differences in Advice Taking: A Meta-Analytic Path Model. Society of Judgement and Decision Making, New York, NY.

Himmelstein, M. (2022). Decline, adopt or compromise? A dual hurdle model for advice utilization. Journal of Mathematical Psychology, 110, 102695. https://doi.org/10.1016/j.jmp.2022.102695


# Thank You!




```{r}
#| fig-align: center
#| fig-height: 7
#| fig-width: 13
#| eval: false

top_plts <- study.level_predictors |>
  head(3) |>
  map(~ (readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis More Predictors",
                           paste0(.x, "_plt_small.rds"))) & 
            theme(geom = element_geom(ink = "#597993", paper = "#a2b9cd"))))

bottom_plts <- study.level_predictors |>
  tail(2) |>
  map(~ (readRDS(here::here("Figures", "Tri-Hurdle Metaanalysis More Predictors",
                           paste0(.x, "_plt_small.rds"))) & 
            theme(geom = element_geom(ink = "#597993", paper = "#a2b9cd"))))

wrap_plots(top_plts, ncol = 3) / 
  wrap_plots(plot_spacer(), wrap_plots(bottom_plts, ncol = 2), plot_spacer(),
             widths = c(0.25, 1, 0.25))
```




