---
title: "When and Why Advice-Taking Strategies Differ:"
subtitle: "A Dual-Hurdle-Based Meta-Analysis"
format: 
  revealjs:
    css: styles.css
width: 1200
author: "Jessica Helmer | TeaP | 2026-03-16"
html-table-processing: none
show-slide-number: all
---

## Advice-Taking

```{r}
#| include: false
library(tidyverse)
library(gt)
library(patchwork)

dat <- readRDS(here::here("..", "Data Clean", "trihurdle-metaanalysis-more-predictors_dat.rds"))
study_names <- readRDS(here::here("..", "Data Clean", "studynames.rds"))

main_color <- "steelblue"
golden_ratio <- (1 + sqrt(5)) / 2

set_theme(theme_classic(base_size = 24))
theme_update(aspect.ratio = golden_ratio ^ -1,
             axis.line.y = element_blank(),
             axis.ticks.y = element_blank(),
             axis.text.x = element_text(size = 30))
```


Studies on advice taking often use a Judge Advisor System setup

<div class="subpoint">○ Judge makes initial judgement</div>

<div class="subpoint">○ Judge receives advice</div>

<div class="subpoint">○ Judge makes final judgement</div>

Outcome metric commonly "Weight of Advice" (WOA)

<!-- Outcome metric of interest is how much someone changed their prediction relative to the advice -->

<br>

:::: {.columns}
::: {.column width="50%"}
$$ WOA = \frac{\text{initial} - \text{final}}{\text{initial} - \text{advice}} $$
:::

::: {.column width="50%"}
```{r}
#| layout-align: center
#| layout-valign: center
tibble(woa = c("$WOA = 0$", "$WOA = 1$", "$0 < WOA < 1$"),
        res = c("Decline", "Adopt", "Compromise")) |>
  gt() |>
  fmt_markdown() |>
  tab_options(column_labels.hidden = T) |>
  opt_table_font(size = 30) |>
  opt_horizontal_padding(3)
```
:::
::::


## Weight of Advice

WOA continuous metric, but often empirically see concentrations of values at 0 and 1

```{r}
#| fig-align: center
#| fig-height: 4
dat |>
  filter(study %in% c(1, 5, 8)) |>
  mutate(study = as.character(study)) |>
  left_join(study_names, by = "study") |>
  mutate(studyname = str_replace_all(studyname, "_", " ")) |>
  ggplot(aes(x = woa_winsor, y = after_stat(density))) +
  geom_histogram(color = NA, fill = main_color, binwidth = 0.08) +
  scale_x_continuous(breaks = c(0, 0.5, 1), labels = c("0", ".5", "1")) +
  scale_y_continuous(limits = c(0, 6.7), breaks = c(0, 6.7), labels = NULL) +
  guides(x = guide_axis(cap = "both")) +
  labs(y = NULL, x = "WOA") +
  facet_wrap(~ studyname, scales = "free_y") +
  theme(panel.spacing = unit(40, "pt"),
             axis.text.x = element_text(size = 20))
```

Typically are modeled with standard linear regression procedures

<div class="subpoint">○ Assume normally distributed residuals!</div>

::: footer
(Himmelstein, 2023)
:::


## Dual-Hurdle Model

Separates outcome into two phases:

:::: {.columns}
::: {.column width="50%"}
**Discrete**

Decline, adopt, or compromise?

Categorical logit: estimates probability of each choice

```{r}
tibble(choice = factor(c("Decline", "Compromise", "Adopt"), 
                       levels = c("Decline", "Compromise", "Adopt"), ordered = T),
       p = c(0.45, 0.25, 0.30)) |>
  ggplot(aes(x = choice, y = p)) +
  geom_col(fill = main_color) +
  scale_y_continuous(NULL, limits = c(0, .46), breaks = NULL) +
  guides(x = guide_axis(NULL, cap = "both"))
```
:::

::: {.column width="50%"}
**Continuous**

If compromise, how much?

Beta distribution: estimates continuous WOA between 0 and 1

```{r}
tibble(woa = rbeta(10000, 1.5, 3)) |>
  ggplot(aes(x = woa)) +
  geom_density(fill = main_color, color = NA, adjust = 3) +
  scale_y_continuous(NULL, limits = c(0, 1.8), breaks = NULL) +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1), labels = c("0", ".5", "1")) +
  guides(x = guide_axis(NULL, cap = "both")) 
```
:::
::::

::: footer
(Himmelstein, 2023)
:::


## Dual-Hurdle Model

:::: {.columns}
::: {.column width="55%"}
Makes WOA distributions like...

```{r}
#| fig-height: 8
d <- rbind(tibble(pc1 = .25,
        pc2 = .45,
        pc3 = .30,
        a = 1.5,
        b = 3) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 1),
   tibble(pc1 = .4,
        pc2 = .5,
        pc3 = .1,
        a = 2,
        b = 4) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 2),
   tibble(pc1 = .7,
        pc2 = .15,
        pc3 = .15,
        a = 3,
        b = 4) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 3),
   tibble(pc1 = .2,
        pc2 = .2,
        pc3 = .6,
        a = 5,
        b = 6) |>
   uncount(1000) |>
   mutate(woa = pmap_dbl(list(pc1, pc2, pc3, a, b), \(pc1, pc2, pc3, a, b) {
     data.frame(res = rmultinom(1, 1, c(pc1, pc2, pc3)),
                woa = c(rbeta(1, a, b), 0, 1)) |>
       filter(res == 1) |>
       pull(woa)}),
     version = 4))

d |>
  ggplot(aes(x = woa)) +
  geom_density(fill = main_color, color = NA, adjust = 1.1) +
  scale_y_continuous(NULL, breaks = NULL) +
  scale_x_continuous(limits = c(0, 1), breaks = c(0, .5, 1),
                     labels = c("0", ".5", "1")) +
  facet_wrap(~ version, ncol = 2, scales = "free_y", axes = "all_x") + 
  guides(x = guide_axis("WOA", cap = "both")) +
  theme(axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        axis.text.x = element_text(size = 25),
        plot.margin = unit(rep(-1, 4), "pt"),
        strip.background = element_blank(),
        strip.text.x = element_blank())
```
:::

::: {.column width="45%"}
Can examine effects of predictors for both the categorical and continuous pieces

Captures variation at the item- and person-level

Present study: extend dual-hurdle model to meta-analytic context

<div class="subpoint">○ Account for study-level variation</div>
:::
::::

::: footer
(Himmelstein, 2023)
:::


## Present Study

Advice taking data from N studies collected in 202X. 

[Details on data collection from Santi's email]

Study-level predictors like student judges, expert advisor, if question was about future events


## Unconditional Model

[Need]


## Tri-Hurdle Model??

Added an extra categorical choice

<div class="subpoint">○ Decline, Adopt, **Midpoint**, Compromise</div>

[Unconditional trihurdle]


## Conditional Model

[Results lol]


## Discussion

[Stuff stuff stuff]


## References


# Thank You!







